biclaR: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools

### 2023-11-27 paper/PaperCEUS/PaperCEUS.Rmd
paper values update + MET

---
title: "biclaR: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 10% of trips could be made by bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 6,000 tons/day, with benefits over 10 years of at least €230 million. For the PT leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

<!-- TODO: -->
<!-- check if HEAT estimates remain the same (the tool is not working atm) -->
<!-- should we ignore the social benefits component? and only estimate and add the environmental ones? or estimate the car to PT transfer Social benefits? maybe to the next paper? -->
<!-- should we keep both direct and safe routing is is it just noise? -->

@@

<!-- **full paper**: 4-6 pages in length (typically up to 3,000 words). we have 2849 words before references, or 2278 words without tables -->

@@

<!-- The quietness level^[see https://www.cyclestreets.net/help/journey/howitworks/#quietness] was another indicator estimated by [CycleStreets](https://www.cyclestreets.org/) for each route section, based on road network variables xisting on OpenStreetMap labels, such as the number of lanes, maximum allowed speed, trees, or road hierarchy. Values vary on a scale of 0 to 100, where zero corresponds to the least safe level for cycling and 100 corresponds to the safest and quietest level, usually corresponding to sections that already include segregated cycling infrastructure away from motorized traffic.  I THINK THIS IS NOT RELEVANT FOR THIS SHORT PAPER, AS IT IS MOST USEFUL FOR VIZUALIZATION PURPOSES ON BICLAR TOOL-->

@@

The routes were then overlaid and aggregated by segments, using [`stplanr overline()` R function](https://docs.ropensci.org/stplanr/reference/overline.html).
<!-- For example, the estimated volumes of bicycle trips and car transfers were aggregated, and the average car speed and quietness level, weighted by the distance of the segments, were calculated. -->

@@

The intermodality scenario considers trips combining PT and cycling for the first and last legs.
In a conservative approach, we have restricted our analysis to the first and last legs with a combined length of up to 5 km (for instance: 1 km from origin to interface A plus 4 km from interface B to destination) or up to 25 minutes on bike.
Furthermore, we have imposed restrictions on PT usage to include only trips with no PT transfers, and up to 2 hours (120 min).
Additionally, we have only included PT modes that can easily accommodate bicycles, such as trains, ferries, trams, and inter-municipal bus lines equipped with bike racks (Figure \ref{fig:map1}).

```{r map1, out.width="60%", fig.cap="Interfaces and lines considered, by transport mode, in the Lisbon metropolitan area", fig.align='center'}
knitr::include_graphics("img/map1.png", error = FALSE)
```

@@

For the cycling legs of the journey (first and last legs), socio-environmental impacts were estimated, using the HEAT for Cycling tool v5.2 [@HEAT] from the World Health Organization, and the [`HEATaaS` R package](https://github.com/HEAT-WHO/HEAT_heatr_api)[^5].
The HEAT tool provided estimates on the shifting from car to cycling for a short term time horizon (i.e., one year) and the long term (i.e., ten years).
We considered two dimensions: *social* — including the physical activity of cyclists, air pollution exposure, and road casualties; *environmental* — including CO~2~eq emissions and other pollutants.

@@

<!-- For the full paper, consider refer the input HEAT values (tabela 6) -->
<!-- Explain why the use of the HaaS tool instead of the website (3 scenarios, 18 municipalities, 2 routing profiles) -->
<!-- Also consider put the FE formula and emmission factor values - see report -->

For the second leg of the journey, we estimate the additional environmental impacts of shifting car trips to PT (between the PT interfaces).

@@

summaryall = summary0 |> 
  left_join(summary22) |> 
  select(Target, Routing, CO2eq, CO2eq_TP, Environmenal_only, Environmenal10_only, Social, Social10, Value) |> 
    mutate(Social = Social/1000,
           Social10 = Social10/1000,
         Environmenal_only = Environmenal_only/1000,
         Environmenal10_only = Environmenal10_only/1000) |> 
  mutate(CO2eqT = CO2eq + CO2eq_TP,
         value1 = Environmenal_only + Value + Social, # this is CORRECT
         value10 = Environmenal10_only + Value*10*rate + Social10) |> # this is CORRECT
  select(Target, Routing, CO2eqT, value1, value10)

@@

<!-- FOR OTHER PAPER: separate the Socio impacts from the Environmental ones? It shouldn't be ok to sum them -->
Shifting from car to cycling + PT can reduce annual CO~2~eq emissions by 14,000 to 36,000 tons per year. 
The 10-year socio-environmental benefits account for €235 million to €620 million, depending on the cycling targets.

The social impacts represent 98% of the socio-environmental benefits (in value) from replacing car trips to bicycle in first-and-last legs. For the PT segment, we did not estimate the social impacts from substituting car trips, although its health benefits would not be as high as shifting to cycling.<!-- We should estimate the social impacts from replacing Car to PT in the second leg, for the other paper --> <!--Overall there was very little evidence available on mode switch to public transport. -->

The emissions of CO~2~eq that are avoided during both the initial and final journey segments account for about 75% of the emissions avoided during the PT segment. This finding, while expected -- due the zero cycling emissions, should not be overlooked when promoting the PT use.
Improving the safe accessibility to PT interfaces to cyclists and providing bicycle-friendly amenities such as parking facilities can potentially lead to a higher reduction in CO~2~eq emissions, compared to a scenario where individuals shift from car travel to car + PT combination.

Our findings suggest that cycling and PT *in combination* could viably replace 10% of current LMA trips, with an additional 6% of PT journeys prone to further substitution. 
<!-- repeated? not correct -->

There are also social benefits form shifting car trips to PT, nevertheless those were not estimated in this research.
One of the main socioenvironmetal benefits, valued after monetization, comes from the increase in physical activity [@Felix2023ES].
The literature shows that the Metabolic Equivalent Tasks (MET) for "riding in a bus or a train" is 1.3 plus the "walking for transportation" is 3.5, while driving a car is 2.5. see https://golf.procon.org/met-values-for-800-activities/
The difference is not that much between them.
Future works should also encompass the estimation of the social impacts for the PT leg of the journey, shifting from car.

>>>>

---
title: "biclaR: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 10% of trips could be made by bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the PT leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

<!-- TODO: -->
<!-- should we ignore the social benefits component? and only estimate and add the environmental ones? or estimate the car to PT transfer Social benefits? maybe to the next paper? -->
<!-- should we keep both direct and safe routing is is it just noise? -->

@@

@@

@@

The routes were then overlaid and aggregated by segments, using [`stplanr overline()` R function](https://docs.ropensci.org/stplanr/reference/overline.html).

@@

The intermodality scenario considers trips combining PT and cycling for the first and last legs.
In a conservative approach, we have restricted our analysis to the first and last legs with a combined length of up to 5 km (for instance: 1 km from origin to interface _A_ plus 4 km from interface _B_ to destination) or up to 25 minutes on bike.
Furthermore, we have imposed restrictions on PT usage to include only trips with no PT transfers, and up to 2 hours (120 min).
Additionally, we have only included PT modes that can easily accommodate bicycles, such as trains, ferries, trams, and inter-municipal bus lines equipped with bike racks (Figure \ref{fig:map1}).

```{r map1, out.width="60%", fig.cap="Interfaces and lines considered, by transport mode, in the Lisbon metropolitan area.", fig.align='center'}
knitr::include_graphics("img/map1.png", error = FALSE)
```

@@

For the *cycling legs of the journey* (first and last legs), socio-environmental impacts were estimated, using the HEAT for Cycling tool v5.2 [@HEAT] from the World Health Organization, and the [`HEATaaS` R package](https://github.com/HEAT-WHO/HEAT_heatr_api)[^5]. 
The use of this package made it possible to run multiple scenarios with few changes in input values, making the interaction with HEAT more reliable when reproducing runs.    
The HEAT tool provided estimates on the shifting from car to cycling for a short term time horizon (i.e., one year) and the long term (i.e., ten years). It estimates the differences between two considered scenarios. In this case: one baseline scenario, with data from the mobility survey, and one cycling potential scenario in which targets of 4% and 10% of cycling levels were achieved, transferred from car trips.
We considered two dimensions: *social* — including the physical activity, air pollution exposure, and road casualties; and *environmental* — including CO~2~eq emissions and other pollutants.

@@

<!-- For the full paper, consider refer the input HEAT values (tabela 6) -->

<!-- include a map or image with the legs and all trip? by mode  -->

<!-- Also consider put the FE formula and emmission factor values - see report -->

For the *second leg of the journey*, we estimate the environmental impacts of shifting car trips to PT (between the PT interfaces).

@@

summaryall = summary0 |> 
  left_join(summary22) |> 
  select(Target, Routing, CO2eq, CO2eq_TP, Environmenal_only, Environmenal10_only, Social, Social10, Value) |> 
    mutate(Social = Social/1000,
           Social10 = Social10/1000,
         Environmenal_only = Environmenal_only/1000,
         Environmenal10_only = Environmenal10_only/1000) |> 
  mutate(CO2eqT = CO2eq + CO2eq_TP,
         value1 = Environmenal_only + Value + Social, # this is CORRECT
         value10 = Environmenal10_only + Value*10*rate + Social10) |> # this is CORRECT
  select(Target, Routing, CO2eqT, value1, value10) |> 
  arrange(desc(Target))

@@

<!-- FOR OTHER PAPER: separate the Socio impacts from the Environmental ones? -->
Shifting from car to cycling + PT can reduce annual CO~2~eq emissions by 14,000 to 36,000 tons per year. 
The 10-year socio-environmental benefits account for €125 million to €325 million, depending on the cycling targets.

The environmental impacts represent less than 2% of the socio-environmental benefits (in value) from replacing car trips to bicycle in first-and-last legs. 
For the PT segment, we did not estimate the social impacts from substituting car trips. 
One of the main socio-environmetal benefits, valued after monetization, comes from the increase in physical activity [@Felix2023ES].
Although are also social benefits form shifting car trips to PT, its health benefits would not be as high as shifting to cycling. 
The literature shows that the Metabolic Equivalent Tasks (MET) for "riding in a bus or a train" is 1.3 plus the "walking for transportation" as 3.5, while "driving a car" is 2.5 [@MET2011].
The difference between these activities is not very obvious when compared to cycling, whose equivalent MET is between 6.8 to 7.5.
Nevertheless, future works should also encompass the estimation of the social impacts for the PT leg of the journey, shifting from car.
 <!--Overall there was very little evidence available on mode switch to public transport. -->

The emissions of CO~2~eq that are avoided during both the initial and final journey segments account for about 74% of the emissions avoided during the PT segment. This finding, while expected -- due the zero cycling emissions, should not be overlooked when promoting the PT use.
Improving the safe accessibility to PT interfaces to cyclists and providing bicycle-friendly amenities such as parking facilities can potentially lead to a higher reduction in CO~2~eq emissions, compared to a scenario where individuals shift from car travel to car + PT combination.

Our findings suggest that cycling and PT *in combination* could viably replace 10% of current LMA trips, with an additional 6% of PT journeys prone to further substitution. 
<!-- repeated? -->

### 2023-11-27 paper/PaperCEUS/PaperCEUS.tex
paper values update + MET

  \begin{abstract}
  A high proportion of car trips can be replaced by a combination of
  public transit and cycling for the first-and-last mile. This paper
  estimates the potential for cycling combined with public transit (PT)
  as a substitute for car trips in the Lisbon metropolitan area and
  assesses its socio-environmental impacts using open data and open
  source tools. A decision support tool that facilitates the design and
  development of a metropolitan cycling network was developed
  (\emph{biclaR}). The social and environmental impacts were assessed
  using the \emph{HEAT for Cycling} and the \emph{HEAT as a Service}
  tools. The impacts of shifting car trips to PT were also estimated and
  monetized. The results indicate that 10\% of trips could be made by
  bicycle + PT combination. Shifting to cycling for the first-and-last
  mile stages can reduce annual CO\textsubscript{2}eq emissions from
  6,000 tons/day, with benefits over 10 years of at least €230 million.
  For the PT leg, the transfer from car avoids of up to 20,500 tons of
  CO\textsubscript{2}eq emissions per year. This evidence can support
  policymakers to prioritize interventions that reduce the reliance on
  private motor vehicles.\\
  \end{abstract}
    \begin{keyword}
    Active transport \sep Intermodality \sep First and last
mile \sep Health economic assessment \sep Environmental impacts \sep 
    Open data and methods
  \end{keyword}

@@

The intermodality scenario considers trips combining PT and cycling for
the first and last legs. In a conservative approach, we have restricted
our analysis to the first and last legs with a combined length of up to
5 km (for instance: 1 km from origin to interface A plus 4 km from
interface B to destination) or up to 25 minutes on bike. Furthermore, we
have imposed restrictions on PT usage to include only trips with no PT
transfers, and up to 2 hours (120 min). Additionally, we have only
included PT modes that can easily accommodate bicycles, such as trains,
ferries, trams, and inter-municipal bus lines equipped with bike racks
(Figure \ref{fig:map1}).

@@

\caption{Interfaces and lines considered, by transport mode, in the Lisbon metropolitan area}\label{fig:map1}
\end{figure}

@@

For the cycling legs of the journey (first and last legs),
socio-environmental impacts were estimated, using the HEAT for Cycling
tool v5.2 (Kahlmeier et al., 2017) from the World Health Organization,
and the
\href{https://github.com/HEAT-WHO/HEAT_heatr_api}{\texttt{HEATaaS} R
package}\footnote{\texttt{HEATaaS} is under development. For more
  information contact
  \href{https://heatwalkingcycling.org}{heatwalkingcycling.org}.}. The
HEAT tool provided estimates on the shifting from car to cycling for a
short term time horizon (i.e., one year) and the long term (i.e., ten
years). We considered two dimensions: \emph{social} --- including the
physical activity of cyclists, air pollution exposure, and road
casualties; \emph{environmental} --- including CO\textsubscript{2}eq
emissions and other pollutants.

For the second leg of the journey, we estimate the additional
environmental impacts of shifting car trips to PT (between the PT
interfaces).

@@

Table \ref{tab:summary22} presents the results of the avoided emissions
and its monetization for the second leg of the journey, by replacing car
trips with potential TP trips. Regarding the PT segment, the shift from
private car would lead to the mitigation of CO\textsubscript{2}
equivalent emissions to 8,500 to 20,800 tons annually, valued in €1.4
million to €3.5 million yearly.

@@

\caption{\label{tab:summaryall}\label{summaryall}Summary of the avoided CO2eq emmissions (ton/year) and the estimated social and environmental benefits (monetized in thousand €) by replacing car trips with cycling in combination with PT.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}}
\toprule
Target & Routing & Avoided CO2eq (tons) & SE Benefits 1yr (k€) & SE Benefits 10yrs (k€)\\
\midrule
4\% & safe & 11 551 & 14 380 & 127 534\\
10\% & safe & 28 217 & 36 861 & 325 814\\
4\% & direct & 11 706 & 14 135 & 125 016\\
10\% & direct & 28 487 & 35 905 & 318 062\\
\bottomrule
\end{tabular}
\end{table}

Shifting from car to cycling + PT can reduce annual
CO\textsubscript{2}eq emissions by 14,000 to 36,000 tons per year. The
10-year socio-environmental benefits account for €235 million to €620
million, depending on the cycling targets.

The social impacts represent 98\% of the socio-environmental benefits
(in value) from replacing car trips to bicycle in first-and-last legs.
For the PT segment, we did not estimate the social impacts from
substituting car trips, although its health benefits would not be as
high as shifting to cycling.

The emissions of CO\textsubscript{2}eq that are avoided during both the
initial and final journey segments account for about 75\% of the
emissions avoided during the PT segment. This finding, while expected --
due the zero cycling emissions, should not be overlooked when promoting
the PT use. Improving the safe accessibility to PT interfaces to
cyclists and providing bicycle-friendly amenities such as parking
facilities can potentially lead to a higher reduction in
CO\textsubscript{2}eq emissions, compared to a scenario where
individuals shift from car travel to car + PT combination.

@@

There are also social benefits form shifting car trips to PT,
nevertheless those were not estimated in this research. One of the main
socioenvironmetal benefits, valued after monetization, comes from the
increase in physical activity (Félix et al., 2023). The literature shows
that the Metabolic Equivalent Tasks (MET) for ``riding in a bus or a
train'' is 1.3 plus the ``walking for transportation'' is 3.5, while
driving a car is 2.5. see
https://golf.procon.org/met-values-for-800-activities/ The difference is
not that much between them. Future works should also encompass the
estimation of the social impacts for the PT leg of the journey, shifting
from car.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-bickel2006}{}}%
Bickel, P., Friedrich, R., Burgess, A., Fagiani, P., Hunt, A., Jong, G.
D., \& Laird, J. (2006). \emph{{HEATCO - Developing Harmonised European
Approaches for Transport Costing and Project Assessment. Deliverable 5,
Proposal for Harmonised Guidelines.}}
\url{https://trimis.ec.europa.eu/sites/default/files/project/documents/20130122_113653_88902_HEATCO_D5_summary.pdf}

@@

\leavevmode\vadjust pre{\hypertarget{ref-goodman2019}{}}%
Goodman, A., Rojas, I. F., Woodcock, J., Aldred, R., Berkoff, N.,
Morgan, M., Abbas, A., \& Lovelace, R. (2019). Scenarios of cycling to
school in england, and associated health and carbon impacts: Application
of the {`}propensity to cycle tool{'}. \emph{Journal of Transport and
Health}, \emph{12}, 263--278.
\url{https://doi.org/10.1016/j.jth.2019.01.008}

>>>>

  \begin{abstract}
  A high proportion of car trips can be replaced by a combination of
  public transit and cycling for the first-and-last mile. This paper
  estimates the potential for cycling combined with public transit (PT)
  as a substitute for car trips in the Lisbon metropolitan area and
  assesses its socio-environmental impacts using open data and open
  source tools. A decision support tool that facilitates the design and
  development of a metropolitan cycling network was developed
  (\emph{biclaR}). The social and environmental impacts were assessed
  using the \emph{HEAT for Cycling} and the \emph{HEAT as a Service}
  tools. The impacts of shifting car trips to PT were also estimated and
  monetized. The results indicate that 10\% of trips could be made by
  bicycle + PT combination. Shifting to cycling for the first-and-last
  mile stages can reduce annual CO\textsubscript{2}eq emissions from
  3,000 to 7,500 tons/day, while for the PT leg, the transfer from car
  avoids of up to 20,500 tons of CO\textsubscript{2}eq emissions per
  year. The estimated socio-environmental benefits are of €125 million
  to €325 million over 10 years. This evidence can support policymakers
  to prioritize interventions that reduce the reliance on private motor
  vehicles.\\
  \end{abstract}
    \begin{keyword}
    Active transport \sep Intermodality \sep First and last
mile \sep Health economic assessment \sep Environmental impacts \sep 
    Open data and methods
  \end{keyword}

@@

The intermodality scenario considers trips combining PT and cycling for
the first and last legs. In a conservative approach, we have restricted
our analysis to the first and last legs with a combined length of up to
5 km (for instance: 1 km from origin to interface \emph{A} plus 4 km
from interface \emph{B} to destination) or up to 25 minutes on bike.
Furthermore, we have imposed restrictions on PT usage to include only
trips with no PT transfers, and up to 2 hours (120 min). Additionally,
we have only included PT modes that can easily accommodate bicycles,
such as trains, ferries, trams, and inter-municipal bus lines equipped
with bike racks (Figure \ref{fig:map1}).

@@

\caption{Interfaces and lines considered, by transport mode, in the Lisbon metropolitan area.}\label{fig:map1}
\end{figure}

@@

For the \emph{cycling legs of the journey} (first and last legs),
socio-environmental impacts were estimated, using the HEAT for Cycling
tool v5.2 (Kahlmeier et al., 2017) from the World Health Organization,
and the
\href{https://github.com/HEAT-WHO/HEAT_heatr_api}{\texttt{HEATaaS} R
package}\footnote{\texttt{HEATaaS} is under development. For more
  information contact
  \href{https://heatwalkingcycling.org}{heatwalkingcycling.org}.}. The
use of this package made it possible to run multiple scenarios with few
changes in input values, making the interaction with HEAT more reliable
when reproducing runs.\\
The HEAT tool provided estimates on the shifting from car to cycling for
a short term time horizon (i.e., one year) and the long term (i.e., ten
years). It estimates the differences between two considered scenarios.
In this case: one baseline scenario, with data from the mobility survey,
and one cycling potential scenario in which targets of 4\% and 10\% of
cycling levels were achieved, transferred from car trips. We considered
two dimensions: \emph{social} --- including the physical activity, air
pollution exposure, and road casualties; and \emph{environmental} ---
including CO\textsubscript{2}eq emissions and other pollutants.

For the \emph{second leg of the journey}, we estimate the environmental
impacts of shifting car trips to PT (between the PT interfaces).

@@

Table \ref{tab:summary22} presents the results of the avoided emissions
and its monetization for the second leg of the journey, by replacing car
trips with potential TP trips.\\
Regarding the PT segment, the shift from private car would lead to the
mitigation of CO\textsubscript{2} equivalent emissions to 8,500 to
20,800 tons annually, valued in €1.4 million to €3.5 million yearly.

@@

\caption{\label{tab:summaryall}\label{summaryall}Summary of the avoided CO2eq emmissions (ton/year) and the estimated social and environmental benefits (monetized in thousand €) by replacing car trips with cycling in combination with PT.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}}
\toprule
Target & Routing & Avoided CO2eq (tons) & SE Benefits 1yr (k€) & SE Benefits 10yrs (k€)\\
\midrule
4\% & safe & 11 551 & 14 380 & 127 534\\
4\% & direct & 11 706 & 14 135 & 125 016\\
10\% & safe & 28 217 & 36 861 & 325 814\\
10\% & direct & 28 487 & 35 905 & 318 062\\
\bottomrule
\end{tabular}
\end{table}

Shifting from car to cycling + PT can reduce annual
CO\textsubscript{2}eq emissions by 14,000 to 36,000 tons per year. The
10-year socio-environmental benefits account for €125 million to €325
million, depending on the cycling targets.

The environmental impacts represent less than 2\% of the
socio-environmental benefits (in value) from replacing car trips to
bicycle in first-and-last legs. For the PT segment, we did not estimate
the social impacts from substituting car trips. One of the main
socio-environmetal benefits, valued after monetization, comes from the
increase in physical activity (Félix et al., 2023). Although are also
social benefits form shifting car trips to PT, its health benefits would
not be as high as shifting to cycling. The literature shows that the
Metabolic Equivalent Tasks (MET) for ``riding in a bus or a train'' is
1.3 plus the ``walking for transportation'' as 3.5, while ``driving a
car'' is 2.5 (Ainsworth et al., 2011). The difference between these
activities is not very obvious when compared to cycling, whose
equivalent MET is between 6.8 to 7.5. Nevertheless, future works should
also encompass the estimation of the social impacts for the PT leg of
the journey, shifting from car.

The emissions of CO\textsubscript{2}eq that are avoided during both the
initial and final journey segments account for about 74\% of the
emissions avoided during the PT segment. This finding, while expected --
due the zero cycling emissions, should not be overlooked when promoting
the PT use. Improving the safe accessibility to PT interfaces to
cyclists and providing bicycle-friendly amenities such as parking
facilities can potentially lead to a higher reduction in
CO\textsubscript{2}eq emissions, compared to a scenario where
individuals shift from car travel to car + PT combination.

@@

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-MET2011}{}}%
Ainsworth, B. E., Haskell, W. L., Herrmann, S. D., Meckes, N., Bassett
Jr, D. R., Tudor-Locke, C., Greer, J. L., Vezina, J., Whitt-Glover, M.
C., \& Leon, A. S. (2011). {2011 Compendium of Physical Activities: A
second update of codes and MET values}. \emph{Medicine \& Science in
Sports \& Exercise}, \emph{43}(8), 1575--1581.
\url{https://doi.org/10.1249/MSS.0b013e31821ece12}

\leavevmode\vadjust pre{\hypertarget{ref-bickel2006}{}}%
Bickel, P., Friedrich, R., Burgess, A., Fagiani, P., Hunt, A., Jong, G.
D., \& Laird, J. (2006). \emph{{HEATCO - Developing Harmonised European
Approaches for Transport Costing and Project Assessment. Deliverable 5,
Proposal for Harmonised Guidelines.}}
\url{https://trimis.ec.europa.eu/sites/default/files/project/documents/20130122_113653_88902_HEATCO_D5_summary.pdf}

@@

\leavevmode\vadjust pre{\hypertarget{ref-goodman2019}{}}%
Goodman, A., Rojas, I. F., Woodcock, J., Aldred, R., Berkoff, N.,
Morgan, M., Abbas, A., \& Lovelace, R. (2019). Scenarios of cycling to
school in england, and associated health and carbon impacts: Application
of the {`}propensity to cycle tool{'}. \emph{Journal of Transport \&
Health}, \emph{12}, 263--278.
\url{https://doi.org/10.1016/j.jth.2019.01.008}
